### 2021-10-30 code/processing_code/Cleaning_Reports.rmd
10 Clean up
Cleaning up and adding notations, making data connections more reproducible.

---
title: "Cleaning_Reports"
author: "MYC"
date: "10/29/2021"
output: html_document
---
Trying to determine the best way to read in excel reports for more automated data building.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(lubridate)
library(broom)
```
## Read in Reports 
Reports are generated in .csv outputs.
Would like to be able to read in each individual file from data\raw_data\Reports
```{r Read Reports}

```

## Scrape Data from reports
From read in data frames need to pull out a selection of data from A21:F12 (includes headers)

```{r Scraping Data}

```