### 2021-11-12 code/processing_code/ProcessingScript2.rmd
Clean up and Temp Code
Clean up of unused scripts and removal of knitted items. Temp Code added-- working on doing a completely automated data import from reports.

---
title: "ProcessingScript2"
author: "MYC"
date: "11/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) #for data processing
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(reshape2)

```
Path to data
```{r Data}
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","P2-20211029.rds")
#read and load data. 
P2<- readRDS(data_location)
```

```{r}
summary(P2)

melt(P2, 
     id.vars = c("Detector", "Method", "Tool"),
     measure.vars = c("Input","Ct"))%>%
  summarise(n=n())

```
```{r}
dcast(P2, Method+Tool+Input+Dilution+Ct~Tool)%>%
  ggplot(aes(y=Ct, x=Dilution))+
  geom_jitter(aes(color=Method))
```