### 2021-11-12 code/processing_code/TEMP/Cleaning_Reports.rmd

---
title: "Cleaning_Reports"
author: "MYC"
date: "10/29/2021"
output: html_document
---
Trying to determine the best way to read in excel reports for more automated data building.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(lubridate)
library(fs)
```
## Read in Reports 
Reports are generated in .csv outputs.
Would like to be able to read in each individual file from data\raw_data\Reports
```{r Path to Reports}
#set the directory/file location of reports
data_dir<-(here::here("data","raw_data","Reports"))

#create path for all csv files in directory of reports
csv_files<-fs::dir_ls(data_dir, # shows list of files in directory
                      regexp = "\\.csv$") #select only.csv files 
```

## Read section of reports
From read in data frames need to pull out a selection of data from A21:F12 (includes headers)

```{r Read and select from Reports}

# read a single report in as a data frame
Report01<-read.csv(csv_files[1],nrow=94, skip=28)[,1:6] #range finding phases
Report02<-read.csv(csv_files[2],nrow=94, skip=28)[,1:6]#range finding phases
Report03<-read.csv(csv_files[3],nrow=94, skip=28)[,1:6]#range finding phases
Report04<-read.csv(csv_files[4],nrow=94, skip=28)[,1:6]#range finding phases
Report05<-read.csv(csv_files[5],nrow=94, skip=28)[,1:6]#range finding phases
Report06<-read.csv(csv_files[6],nrow=76, skip=28)[,1:6] #starts Phase 2
Report07<-read.csv(csv_files[7],nrow=76, skip=28)[,1:6] #starts Phase 2
Report08<-read.csv(csv_files[8],nrow=76, skip=28)[,1:6] #starts Phase 2
Report09<-read.csv(csv_files[9],nrow=76, skip=28)[,1:6] #starts Phase 2
Report10<-read.csv(csv_files[10],nrow=76, skip=28)[,1:6] #starts Phase 2
Report11<-read.csv(csv_files[11],nrow=76, skip=28)[,1:6] #starts Phase 2

```

```{r All Data}
#merge all same variables from all reports
AllData<-rbind(Report01,
      Report02,
      Report03,
      Report04,
      Report05,
      Report06,
      Report07,
      Report08,
      Report09,
      Report10,
      Report11
      )
```

```{r P1}
#selective merge of data for part 1
P1<-rbind(Report01,
      Report02,
      Report03,
      Report04,
      Report05
      )
```

```{r P2}
#selective merge of data part 2
P2<-rbind(
      Report06,
      Report07,
      Report08,
      Report09,
      Report10,
      Report11
      )
```